---
title: "data viz draft"
author: "Shijing Zhou"
date: "2/24/2021"
output:
  html_document:
    toc: yes
    toc_float: TRUE
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(colorblindr)
library(viridis)
opts_chunk$set(echo = TRUE)
options(scipen = 999)
nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")
```


# Visualization 1: Squirrel's behavior by age

```{r}
age_squirrels <- nyc_squirrels %>% 
  select(age, running, chasing, climbing, eating, foraging) %>% 
  group_by(age) %>% 
  summarize(.groups = "keep", # Adding .groups = "keep" removes the warning that it's only keeping the final group you list - if you link several group_by() functions with %>%, .groups = "keep" keeps all those groups for you.
    Running = sum(running, na.rm = TRUE), 
            Chasing = sum(chasing, na.rm = TRUE), 
            Climbing = sum(climbing, na.rm = TRUE), 
            Eating = sum(eating, na.rm = TRUE), 
            Foraging = sum(foraging, na.rm = TRUE)) %>% 
  subset(age == "Adult" | age == "Juvenile") %>% # I really like this use of subset() - I haven't used it before and like it alot!
  pivot_longer(!age, names_to = "behavior", values_to = "count")

age_squirrels %>% 
  ggplot(aes(age, count, 
             fill = factor(behavior),
             group = factor(behavior))) +
  geom_bar(stat="identity", position="fill", size=0.6) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Age Group",
       y = "Proportions of Behavior",
       fill = "Behaviors",
       title = "Squirrel behaviors in different age groups",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

```
I love this data set and I think this is a delightful vizualization! I wonder if, depending on your communication goal, you may find success flipping it to be something like this:

```{r Zach suggestion for Viz1}
prop_squirrels <- age_squirrels %>% 
  group_by(age) %>% 
  mutate( 
  prop = count/sum(count)
  )

prop_squirrels %>% 
  ggplot(aes(behavior, prop, 
             fill = factor(age),
             group = factor(age))) +
  geom_bar(stat="identity", position="dodge", size=0.6) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Age Group",
       y = "Proportions of Behavior",
       fill = "Behaviors",
       title = "Squirrel behaviors in different age groups",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```


# Visualization 2: Personality and tail position of squirrels 

```{r}
tail_squirrels <- nyc_squirrels %>% 
  select(tail_flags, tail_twitches, approaches, indifferent, runs_from) 

tail_flag <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "FALSE")
tail_twitch <- subset(tail_squirrels, tail_twitches == "TRUE" & tail_flags == "FALSE")
tail_both <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "TRUE")
tail_no_move <- subset(tail_squirrels, tail_flags == "FALSE" & tail_twitches == "FALSE")

tail_flag <- tail_flag %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_flag <- tail_flag %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_flag["tail"] <- "Flagging"


tail_twitch <- tail_twitch %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_twitch <- tail_twitch %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_twitch["tail"] <- "Twitching"


tail_both <- tail_both %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_both <- tail_both %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_both["tail"] <- "Flagging and Twitching"


tail_no_move <- tail_no_move %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_no_move <- tail_no_move %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_no_move["tail"] <- "No Movement"


tail_squirrels <- rbind(tail_flag, tail_twitch, tail_both) #not including no movement because it confuses other group data

#bar chart
# CREATING PROPORTION VARIABLE GROUPED BY THE SQUIRREL PERSONALITY
prop_tail <- tail_squirrels %>% 
  group_by(personality) %>% 
  mutate( 
  prop = count/sum(count)
  )
# BAR CHART
prop_tail %>% 
  ggplot(aes(personality, prop, 
             fill = factor(tail),
             group = factor(tail))) + 
  coord_flip() +
  geom_bar(stat="identity", position="dodge", size=0.6) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Personality",
       y = "Proportions of Squirrel's Personality",
       fill = "Tail Movement",
       title = "Personality and Tail Position of Squirrels",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "bottom")

#pie chart - need more formatting with label
tail_squirrels %>% 
  ggplot(aes(tail, count, 
             fill = factor(personality))) + 
  geom_bar(stat="identity", position="fill", size=0.6) + 
  coord_polar("y", start=0) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_void() + 
  scale_fill_OkabeIto() +
  labs(x = "Tail Movement",
       y = "Proportions of Squirrel's Personality",
       fill = "Personality",
       title = "Personality and Tail Position of Squirrels",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
  facet_wrap(~tail, nrow = 1)
```

This is another delightful dataset and visualization - I had no idea ggplot let you map concentric pie charts! My first comment was that a reader may be confused by what each ring in the pie chart signifies, which I see you've noted in the comment above. 

I played around with the bar chart above and wonder what you think of flipping the grouping variables - again, it all depends on what you're trying to communicate! I think, as a reader, I'm particularly interested in how the different personalities behave, so I wonder if flipping the bar chart to organize by personality types would be helpfuL. I recalculated the proportions to add up to 1 within each squirrel personality, which may skew perceptions of the relative population of "friendly" to "runner" squirrels. 

# Visualization 3: Geographical effect of Squirrels

```{r}
geo_squirrels <- nyc_squirrels %>% 
  group_by(hectare) %>% 
  count() %>%
  separate(hectare, 
           into = c("north-to-south", "east-to-west"),
           sep = "(?<=[0-9])(?=[A-Za-z])")

geo_squirrels %>% 
  ggplot(aes(`east-to-west`,`north-to-south`)) + 
  geom_tile(aes(fill = n),
            color = "white") + 
  coord_fixed() +
  scico::scale_fill_scico(palette = "tokyo") +
  theme_classic() + 
  labs(title = "Squirrel Observations in NYC Central Park",
       subtitle = "Grey area represents geographical areas that with \nno observation of squirrels",
       y = "North to South",
       x = "East to West",
       fill = "Number of Squirrels") + 
 theme(panel.background = element_rect(fill = "light grey"),
       legend.position = "right",
       plot.title = element_text(face = "bold")) 
```

I'm running out of time on my review, but I love this plot! It's delightful! 
The only small tweak that would improve readability that I made was switching the axes so that it matched the layout of Central Park (i.e., so that the y-axis was north-south).
If you want, you could play with the data such that the streets (because NYC is on a grid) match up with your tiles, but I'm unsure how complicated an undertaking that is :-)

Areas of strength: 
1) The map of NYC central park is delightful and very clear!
2) Consistently being color-blind friendly 
3) Great title/subtitles - they're clear and very effective

I learned two things from reading your code - subset() and that you can make concentric pie charts!
