---
title: "NYC Squirrel Census"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: united
    social: menu
    source_code: https://github.com/shijing-z/EDLD652-Final-Project.git
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(knitr)
library(colorblindr)
library(viridis)
opts_chunk$set(echo = FALSE)
options(scipen = 999)
nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")
```

Data Source {.sidebar}
==========
**Data Source**

The data is from the [NYC Squirrel Census](https://www.thesquirrelcensus.com/).


[NYC Squirrel Census Repo from TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29)


Raw data can be found at [NY Data portal](https://data.cityofnewyork.us/Environment/2018-Central-Park-Squirrel-Census-Squirrel-Data/vfnx-vebw).

**Visualization #1**

Squirrel’s behaviors in different age groups of squirrels.

**Visualization #2**

Personality and tail position of squirrels.

**Visualization #3**

Geographical effect of Squirrels.


**Special thanks to my reviewers: **

*Zach*, especially for helpful code suggestions

*Wanjia*

*Joe*

Visualization 1 {data-icon="fa-signal"}
======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Version 1

```{r echo = FALSE}
age_squirrels <- nyc_squirrels %>% 
  select(age, running, chasing, climbing, eating, foraging) %>% 
  group_by(age) %>% 
  summarize(Running = sum(running, na.rm = TRUE), 
            Chasing = sum(chasing, na.rm = TRUE), 
            Climbing = sum(climbing, na.rm = TRUE), 
            Eating = sum(eating, na.rm = TRUE), 
            Foraging = sum(foraging, na.rm = TRUE),
            .groups = "keep") %>% 
  subset(age == "Adult" | age == "Juvenile") %>% 
  pivot_longer(!age, names_to = "behavior", values_to = "count") 

age_squirrels$age <- factor(age_squirrels$age, levels = c("Juvenile", "Adult"))

prop_squirrels <- age_squirrels %>% 
  group_by(age) %>% 
  mutate(prop = count/sum(count))

pd <- position_dodge(1)

prop_squirrels %>% 
  ggplot(aes(behavior, prop, 
             fill = age,
             group = age)) +
  geom_bar(stat = "identity", position = pd, size = 0.6) + 
  geom_text(aes(label=scales::percent(prop, 1)), 
            size = 4.5,
            position = pd,
            vjust = 1.5,
            color = "white") + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Behaviors",
       y = "Proportions of Behavior",
       fill = "Age",
       title = "Squirrel behaviors in different age groups",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold"))
```

### Version 2

```{r echo = FALSE}
age_squirrels <- nyc_squirrels %>% 
  select(age, running, chasing, climbing, eating, foraging) %>% 
  group_by(age) %>% 
  summarize(Running = sum(running, na.rm = TRUE), 
            Chasing = sum(chasing, na.rm = TRUE), 
            Climbing = sum(climbing, na.rm = TRUE), 
            Eating = sum(eating, na.rm = TRUE), 
            Foraging = sum(foraging, na.rm = TRUE),
            .groups = "keep") %>% 
  subset(age == "Adult" | age == "Juvenile") %>% 
  pivot_longer(!age, names_to = "behavior", values_to = "count")


age_squirrels_pct <- age_squirrels %>% 
  group_by(age) %>% 
  mutate(percent = count/sum(count))

age_squirrels_pct <-  age_squirrels_pct %>% 
  group_by(age) %>% 
  mutate(cummulative = cumsum(percent),
         halves = percent/2,
         position = cummulative - halves,
         behavior = forcats::fct_inorder(behavior))

age_squirrels_pct %>% 
  ggplot(aes(age, percent, 
             fill = forcats::fct_rev(behavior),
             group = forcats::fct_rev(behavior))) +
  geom_col(size=0.6) + 
  geom_text(aes(label = paste0(round(percent, 2)*100, "%"),
                y = position), 
            size = 4,
            color = "black") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Age Group",
       y = "Proportions of Behavior",
       fill = "Behaviors",
       title = "Squirrel behaviors in different age groups",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold")) +
  coord_flip()
```


### Draft 

```{r echo = FALSE}
age_squirrels <- nyc_squirrels %>% 
  select(age, running, chasing, climbing, eating, foraging) %>% 
  group_by(age) %>% 
  summarize(Running = sum(running, na.rm = TRUE), 
            Chasing = sum(chasing, na.rm = TRUE), 
            Climbing = sum(climbing, na.rm = TRUE), 
            Eating = sum(eating, na.rm = TRUE), 
            Foraging = sum(foraging, na.rm = TRUE)) %>% 
  subset(age == "Adult" | age == "Juvenile") %>% 
  pivot_longer(!age, names_to = "behavior", values_to = "count")

age_squirrels %>% 
  ggplot(aes(age, count, 
             fill = factor(behavior),
             group = factor(behavior))) +
  geom_bar(stat="identity", position="fill", size=0.6) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Age Group",
       y = "Proportions of Behavior",
       fill = "Behaviors",
       title = "Squirrel behaviors in different age groups",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

Visualization 2 {data-icon="fa-signal"}
======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Final version

```{r echo = FALSE}
tail_squirrels <- nyc_squirrels %>% 
  select(tail_flags, tail_twitches, approaches, indifferent, runs_from) 

tail_flag <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "FALSE")
tail_twitch <- subset(tail_squirrels, tail_twitches == "TRUE" & tail_flags == "FALSE")
tail_both <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "TRUE")
tail_no_move <- subset(tail_squirrels, tail_flags == "FALSE" & tail_twitches == "FALSE")

tail_flag <- tail_flag %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_flag <- tail_flag %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_flag["tail"] <- "Flagging"


tail_twitch <- tail_twitch %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_twitch <- tail_twitch %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_twitch["tail"] <- "Twitching"


tail_both <- tail_both %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_both <- tail_both %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_both["tail"] <- "Flagging and Twitching"


tail_no_move <- tail_no_move %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_no_move <- tail_no_move %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_no_move["tail"] <- "No Movement"


tail_squirrels <- rbind(tail_flag, tail_twitch, tail_both) #not including no movement because it confuses other group data

tail_squirrels$tail <- factor(tail_squirrels$tail, 
                              levels = c("Twitching", "Flagging", "Flagging and Twitching"))

prop_tail <- tail_squirrels %>% 
  group_by(personality) %>% 
  mutate(prop = count/sum(count))

prop_tail <- prop_tail %>% 
  mutate(hjust_amount = ifelse(prop < .05, -0.5, 1.5),
         txt_color = ifelse(prop < .05, "gray20", "white"))

prop_tail %>% 
  ggplot(aes(personality, prop, 
             fill = forcats::fct_rev(tail),
             group = forcats::fct_rev(tail))) + 
  geom_bar(stat="identity", position = pd, size = 0.6) + 
  geom_text(aes(label=scales::percent(prop, 1),
                hjust = hjust_amount,
                color = txt_color), 
            size = 3,
            position = pd) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  scale_color_identity() +
  labs(x = "Personality",
       y = "Proportions of Squirrel's Personality",
       fill = "Tail Movement",
       title = "Personality and Tail Position of Squirrels",
       caption = "Data from NYC Squirrel Census") + 
  coord_flip() +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom") 
```

### Draft

```{r echo = FALSE}
tail_squirrels <- nyc_squirrels %>% 
  select(tail_flags, tail_twitches, approaches, indifferent, runs_from) 

tail_flag <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "FALSE")
tail_twitch <- subset(tail_squirrels, tail_twitches == "TRUE" & tail_flags == "FALSE")
tail_both <- subset(tail_squirrels, tail_flags == "TRUE" & tail_twitches == "TRUE")
tail_no_move <- subset(tail_squirrels, tail_flags == "FALSE" & tail_twitches == "FALSE")

tail_flag <- tail_flag %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_flag <- tail_flag %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_flag["tail"] <- "Flagging"


tail_twitch <- tail_twitch %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_twitch <- tail_twitch %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_twitch["tail"] <- "Twitching"


tail_both <- tail_both %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_both <- tail_both %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_both["tail"] <- "Flagging and Twitching"


tail_no_move <- tail_no_move %>% 
  summarize(Friendly = sum(approaches, na.rm = TRUE),
            Indifferent = sum(indifferent, na.rm = TRUE),
            Runner = sum(runs_from, na.rm = TRUE)) 

tail_no_move <- tail_no_move %>% 
  pivot_longer(cols = 1:3,
               names_to = "personality",
               values_to = "count") 
tail_no_move["tail"] <- "No Movement"


tail_squirrels <- rbind(tail_flag, tail_twitch, tail_both) #not including no movement because it confuses other group data

#bar chart
tail_squirrels %>% 
  ggplot(aes(tail, count, 
             fill = factor(personality),
             group = factor(personality))) + 
  geom_bar(stat="identity", position="fill", size=0.6) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() + 
  scale_fill_OkabeIto() + 
  labs(x = "Tail Movement",
       y = "Proportions of Squirrel's Personality",
       fill = "Personality",
       title = "Personality and Tail Position of Squirrels",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

#pie chart - need more formatting with label
tail_squirrels %>% 
  ggplot(aes(tail, count, 
             fill = factor(personality))) + 
  geom_bar(stat="identity", position="fill", size=0.6) + 
  coord_polar("y", start=0) + 
  scale_y_continuous(expand = c(0,0)) +
  theme_void() + 
  scale_fill_OkabeIto() +
  labs(x = "Tail Movement",
       y = "Proportions of Squirrel's Personality",
       fill = "Personality",
       title = "Personality and Tail Position of Squirrels",
       caption = "Data from NYC Squirrel Census") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

Visualization 3 {data-icon="fa-signal"}
======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Final version

```{r echo = FALSE}
geo_squirrels <- nyc_squirrels %>% 
  group_by(hectare) %>% 
  count() %>%
  separate(hectare, 
           into = c("north-to-south", "east-to-west"),
           sep = "(?<=[0-9])(?=[A-Za-z])")

geo_squirrels %>% 
  ggplot(aes(`north-to-south`, `east-to-west`)) + 
  geom_tile(aes(fill = n),
            color = "white") + 
  coord_fixed() +
  scico::scale_fill_scico(palette = "tokyo") +
  theme_classic() + 
  labs(title = "Squirrel Observations in NYC Central Park",
       subtitle = "Grey area represents geographical areas with no observation of squirrels",
       fill = "Number of Squirrels\n") + 
  xlab(expression(Sourth %->% North)) +
  ylab(expression(East %->% West)) +
  theme(panel.background = element_rect(fill = "light grey"),
       legend.position = "bottom",
       legend.key.width = unit(2, "cm"),
       plot.title = element_text(face = "bold"),
       axis.text.x = element_blank(),
       axis.text.y = element_blank())
```

### Draft

```{r echo = FALSE}
geo_squirrels <- nyc_squirrels %>% 
  group_by(hectare) %>% 
  count() %>%
  separate(hectare, 
           into = c("north-to-south", "east-to-west"),
           sep = "(?<=[0-9])(?=[A-Za-z])")

geo_squirrels %>% 
  ggplot(aes(`north-to-south`, `east-to-west`)) + 
  geom_tile(aes(fill = n),
            color = "white") + 
  coord_fixed() +
  scico::scale_fill_scico(palette = "tokyo") +
  theme_classic() + 
  labs(title = "Squirrel Observations in NYC Central Park",
       subtitle = "Grey area represents geographical areas that with no observation of squirrels",
       y = "East to West",
       x = "North to South",
       fill = "Number of Squirrels") + 
 theme(panel.background = element_rect(fill = "light grey"),
       legend.position = "bottom",
       plot.title = element_text(face = "bold")) 
```